---
title: "model_building_pa_project"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, comment="  ", prompt = TRUE)

#library(tidyverse)
library(mice)
library(dplyr)
library(VIM)
library(ggplot2)
library(lattice)
library(Amelia)
library(caret)
library(doParallel)
library(imputeMissings)
#https://www.kaggle.com/captcalculator/imputing-missing-data-with-the-mice-package-in-r  #great reference

set.seed(12345)
```


note this is the data i was using in winter 2019, however i made a few tuning adjustments in my pipeline while workign on the PA project. I have not updated my origional data since then, and can do this once i have my modelformatting and modeling pipeline in place. Overall, I do not expect to see any noticeable difference in the data.
```{r}
worst2<- read.csv("/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/data/processed/04042019_worstdf_preImp.csv",stringsAsFactors=TRUE, check.names=FALSE)
newagg<- read.csv("/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/data/processed/04042019_newaggdf_preImp.csv",stringsAsFactors=TRUE, check.names=FALSE)
final_pt<- read.csv("/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/data/raw/csv/04042019_final_pt_df2_v.csv",stringsAsFactors=TRUE, check.names=FALSE)

####SHIT, i was using wrong patient list. changed to my most up to date version. the mistake version i used was the patients for the PA project. this had 8,629 icustay id. 
##after updating it now contains 14478 patients.



```

```{r}
dim(final_pt)
final_pt
```


```{r some quick data cleaning and housekeeping}


worst2$first_admit_age[worst2$first_admit_age>90]<- 90 #since the 
colnames(worst2)[colnames(worst2)=="('max', 'sodium')"] <- "maxSodium"
colnames(worst2)[colnames(worst2)=="('min', 'sodium')"] <- "minSodium"
colnames(worst2)[colnames(worst2)=="('max', 'calcium')"] <- "maxCalcium"
colnames(worst2)[colnames(worst2)=="('min', 'calcium')"] <- "minCalcium"
colnames(worst2)[colnames(worst2)=="('max', 'sodium')"] <- "maxSodium"
colnames(worst2)[colnames(worst2)=="('min', 'sodium')"] <- "minSodium"
colnames(worst2)[colnames(worst2)=="('max', 'wbc')"] <- "maxWBC"
colnames(worst2)[colnames(worst2)=="('min', 'wbc')"] <- "minWBC"
colnames(worst2)[colnames(worst2)=="bands"] <- "bandsNum"
colnames(worst2)[colnames(worst2)=="pao2/fio2"] <- "pao2fio2Ratio"


##changing 90 to be max age instead of 300
worst2$first_admit_age[worst2$first_admit_age>90]<- 90

worst2$dopamine<- as.factor(worst2$dopamine == 1)
worst2$dobutamine<- as.factor(worst2$dobutamine == 1)
worst2$vasopressin<- as.factor(worst2$vasopressin == 1)
worst2$epinephrine<- as.factor(worst2$epinephrine == 1)
worst2$norepinephrine<- as.factor(worst2$norepinephrine == 1)
worst2$phenylephrine<- as.factor(worst2$phenylephrine == 1)
worst2$rrt<- as.factor(worst2$rrt == 1)


worst2
final_pt
#final_pt[is.na(final_pt$Gram_pos),'Gram_pos'] <-3
#final_pt<- subset(final_pt, Gram_pos!=2) #REMOVING MYCOBACERIA PATIENTS

#making binary culture variable
#final_pt$iGram_pos<- as.factor(final_pt$Gram_pos==1) #for only culture pos, if gram pos=True, else False

##making 3 level factor variable
#final_pt$iGram_pos<- factor(final_pt$Gram_pos, labels=c('gram_neg','gram_pos','culture_neg'))
#final_pt$iCulture<- as.factor(!is.na(final_pt$Gram_pos)) #a culture being present =True, else False


```

```{r}
##converting bands and pco2 into categorical factors
worst2$ibands<- 'absent'
worst2[which( worst2$bandsNum<=10 & is.na(worst2$bandsNum)==FALSE) , 'ibands'] <- '<10'
worst2[which( worst2$bandsNum >10 & is.na(worst2$bandsNum)==FALSE) , 'ibands'] <- '>10'
summary(as.factor(worst2$ibands))


worst2$ipco2<- 'absent'
worst2[which( worst2$pco2<=50 & is.na(worst2$pco2)==FALSE) , 'ipco2'] <- '<50'
worst2[which( worst2$pco2 >50 & is.na(worst2$pco2)==FALSE) , 'ipco2'] <- '>50'
summary(as.factor(worst2$ipco2))

worst2<- worst2[ , !names(worst2) %in% c('pco2','bandsNum')]

```


```{r}
final_pt %>% 
  select(final_bin) %>% 
  group_by(final_bin) %>% 
  dplyr::summarise(n())

worst2 %>% 
  select(ibands) %>% 
  group_by(ibands) %>% 
  dplyr::summarise(n())

```


```{r merging final clinical data with patient data to ensure all rows have the outcome and sufficient identifier }

#left join worst2 with final_pt outcome and pt id labels
worst2_sid= merge(worst2,final_pt[c('icustay_id','subject_id', 'final_bin')], by='icustay_id', all.x=TRUE)

#  #REMOVING PT WITH MYCOBACTERIA
# worst2_sid<- worst2_sid[worst2_sid$icustay_id %in% final_pt$icustay_id,]

#left join
#worst2_sid$Var.2 <- NULL
worst2_sid<- worst2_sid[,!(names(worst2_sid) %in% c('Var.2'))] #dropping a weird var.2. 
glimpse(worst2_sid)

```


#####filtering to only twoclass:

```{r}
worst2_sid=worst2_sid[worst2_sid$final_bin %in% c("C_neg/A_partial","C_pos/A_full"),]
```


---------------
splitting cohort into two sets, training and validation. note if pt has >1 icustay_id, they should both end up in same cohort. ie they will be split at the subject_id level.
```{r}
summary(worst2_sid$subject_id)
```


```{r splitting cohort into two sets, training and validation.}
set.seed(12345)
index <- createDataPartition(worst2_sid$subject_id, p=0.70, list=FALSE)
trainSet <- worst2_sid[ index,]
testSet <- worst2_sid[-index,]

```

```{r}
colSums(is.na(trainSet))
trainSet

# write.csv(trainSet, file ="/Users/geickelb1/Box/PAII_project/processed_data/pre_imputation_datasets/20190401_train_noimpute.csv", row.names=FALSE)
```



```{r, imputing via median/mode}
median_imputed<- impute(trainSet, method="median/mode")
```

```{r get dataset ready to model} 
#convert outcom to factor again
median_imputed$final_bin<- as.factor(median_imputed$final_bin)

#removing subject id
median_imputed<-median_imputed[, !names(median_imputed)%in% c('subject_id')]
#summary(median_imputed$iGram_pos)
#typeof(median_imputed$iGram_pos)
glimpse(median_imputed)
```

```{r}
write.csv(median_imputed, file ="/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/models/imputation/01042019_median_imputed_train.csv", row.names=FALSE) #"/Users/geickelb1/Documents/GitHub/PAII_mimic_project/processed_data/20190221_median_imputed.csv")

```



### test set

```{r}

median_imputed_test<- impute(testSet, method="median/mode")
#convert outcom to factor again
median_imputed_test$final_bin<- as.factor(median_imputed_test$final_bin)

#removing subject id
median_imputed_test<-median_imputed_test[, !names(median_imputed_test)%in% c('subject_id')]
#summary(median_imputed$iGram_pos)
#typeof(median_imputed$iGram_pos)
glimpse(median_imputed_test)
```



```{r}
write.csv(median_imputed_test, file ="/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/models/imputation/01042019_median_imputed_test.csv", row.names=FALSE)
```


```{r}
preprocessed<- preProcess(median_imputed[,c(2:47)], method = c("center","scale")) #convert to z-scores

preprocessed_imputd<-predict(preprocessed, newdata= median_imputed)
glimpse(preprocessed_imputd)

preprocessed_test<- preProcess(median_imputed_test[,c(2:47)], method = c("center","scale")) #convert to z-scores

preprocessed_imputd_test<-predict(preprocessed_test, newdata= median_imputed_test)

```


```{r}
write.csv(preprocessed_imputd, file ="/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/models/imputation/01042019_median_imputed_train_normalized.csv", row.names=FALSE)

write.csv(preprocessed_imputd_test, file ="/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/models/imputation/01042019_median_imputed_test_normalized.csv", row.names=FALSE)
```






###data dictionary
```{r}
library(ForImp)
missing<- missingness(worst2_sid)
missing_pct<-missing$missing_values_per_variable/ nrow(worst2_sid)
missing_pct<-data.frame(missing_pct)

##grabbing table source data from python output
datadict<- read.csv("/Users/geickelb1/Documents/GitHub/mimiciii-antibiotics-modeling/data/processed/PA_project/merged/25012019_pa_data_dictionary.csv",stringsAsFactors=TRUE, check.names=FALSE)


datadict

missing_pct$label<- rownames(missing_pct)

data.dictionary<-merge(missing_pct,datadict, by='label', all.x=TRUE)

data.dictionary
```

```{r}
write.csv(data.dictionary,"/Users/geickelb1/Documents/GitHub/PAII_mimic_project/model_building/data_dictonary.csv")
#/Users/geickelb1/Documents/GitHub/PAII_mimic_project/model_building/Rplots.pdf
```

